---
title: "Melanoma Sample Integration"
author: "Kari Stromhaug and Giacomo Oliveira"
date: "4/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries and setting path
```{r loading libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(ggpubr)
library(cowplot)
library(Seurat)
library(openxlsx)
library(stringr)
library(clustree)
library(harmony)
library(tidyr)
library(ggplot2)
library(tidyverse)

```


```{r read in data}
setwd("~/Dropbox (Partners HealthCare)/Melanoma P2 Analysis/")
p2.tils = readRDS('patient2/R_output/Seurat_Objects/TILs.Res04.clustered.seurat.20200228.rds')
p6.tils = readRDS('patient6/R_output/Seurat_Objects/TILs.clustered.seurat.20200306.rds')
p11.tils = readRDS('patient11/R_output/Seurat_Objects/TILs.Res03.clustered.seurat.20200309.rds')
p15.tils = readRDS('patient15/R_output/Seurat_Objects/TILs.Res05.clustered.seurat.20200302.rds')

```

```{r}
tils.all <- merge(x = p2.tils, y = c(p6.tils, p11.tils, p15.tils), add.cell.ids = c("p2", "p6", 'p11', 'p15'), project = "Melanoma TCR TILs")

```


```{r}
origs = tils.all@meta.data[,c('orig.ident', 'sample')]; origs$s = rownames(origs)
news = origs %>% separate (s, sep='_', into=c('patient', 'barcode'))
patient = news$patient; names(patient) = rownames(news)
tils.all = AddMetaData(tils.all, metadata=patient, col.name='patient')

```

# Harmony
```{r look at PCs}
tils.harmony <- tils.all %>% NormalizeData(assay = "RNA", normalization.method = "LogNormalize", scale.factor = 1e6)
tils.harmony <- tils.harmony %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000)
tils.harmony <- tils.harmony %>% ScaleData()
tils.harmony <- tils.harmony %>% RunPCA(features = VariableFeatures(tils.harmony), verbose = FALSE)

```


#### Run Harmony
```{r run harmony}
library(harmony)
options(repr.plot.height = 2.5, repr.plot.width = 6)
tils.harmony <- tils.harmony %>% RunHarmony("patient", plot_convergence = TRUE)

options(repr.plot.height = 5, repr.plot.width = 12)
dp1 <- DimPlot(object = tils.harmony, reduction = "harmony", pt.size = .1, group.by = "patient")
v1 <- VlnPlot(object = tils.harmony, features = "harmony_1", group.by = "patient", pt.size = .1)
plot_grid(dp1, v1)
DimPlot(object = tils.harmony, dims=c(1,2), reduction = "harmony", pt.size = .1, group.by = "patient", split.by='patient')

```


#### Cluster again
```{r recluster}

tils.harmony <- tils.harmony %>% 
    RunUMAP(reduction = "harmony", dims=1:50) %>%
    FindNeighbors(reduction = "harmony", dims = 1:50) %>% 
    FindClusters(resolution = 0.5) %>% 
    identity()

DimPlot(tils.harmony, reduction='umap', label=TRUE)
DimPlot(tils.harmony, reduction='umap', split.by='patient')

```


## Run UMAP on tils.harmony cells at different resolutions
```{r res til, message=FALSE, warning=FALSE}
res = c(0.30, 0.35, 0.40, 0.45, 0.50, 0.55, 0.60, 0.65)#, 0.70)#, 0.75, 0.80)#, 0.85, 0.90)

for (i in res) {
  tils.harmony <- FindClusters(tils.harmony, resolution=i)
  tils.harmony@meta.data[,paste0('clusters_', as.character(i))]<-tils.harmony$seurat_clusters
}

clustree(tils.harmony)

```

### Plotting different resolutions
```{r plotting tils.harmony resolutions, message=FALSE, warning=FALSE}
Idents(tils.harmony) <- tils.harmony$clusters_0.3
d1 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.30")

Idents(tils.harmony) <- tils.harmony$clusters_0.35
d2 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.35")

Idents(tils.harmony) <- tils.harmony$clusters_0.4
d3 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.40")

Idents(tils.harmony) <- tils.harmony$clusters_0.45
d4 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.45")

Idents(tils.harmony) <- tils.harmony$clusters_0.5
d5 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.50")

Idents(tils.harmony) <- tils.harmony$clusters_0.55
d6 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.55")

Idents(tils.harmony) <- tils.harmony$clusters_0.6
d7 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.60")

Idents(tils.harmony) <- tils.harmony$clusters_0.65
d8 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.65")

# Idents(tils.harmony) <- tils.harmony$clusters_0.7
# d9 = DimPlot(tils.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.70")

plot_grid(d1, d2, d3, d4)
plot_grid(d5, d6, d7, d8)

Idents(tils.harmony)<-tils.harmony$clusters_0.6
tils.harmony$clusters<-tils.harmony$clusters_0.6

```


### Look at a few features
```{r plotting various things}
p1 = FeaturePlot(tils.harmony, features='cd4.adt')
p2 = FeaturePlot(tils.harmony, features='cd8.adt')
p3 = FeaturePlot(tils.harmony, features='dp.adt')
p4 = FeaturePlot(tils.harmony, features='dn.adt')
plot_grid(p1, p2, p3, p4)

k1 = FeaturePlot(tils.harmony, features='CD19.CITEseq.norm', min.cutoff = 1)
k2 = FeaturePlot(tils.harmony, features='CD3.CITEseq.norm', min.cutoff = 1)
k3 = FeaturePlot(tils.harmony, features='CD11b.CITEseq.norm', min.cutoff = 1)
plot_grid(k1, k2, k3)
FeaturePlot(tils.harmony, feature = "FOXP3")

```


### Renaming Clonotypes
```{r}
meta = tils.harmony@meta.data
are.clones = rownames(meta[!is.na(meta$til.clonotype.family) | !is.na(meta$til.pre.clonotype.family) | !is.na(meta$til.rel.clonotype.family), ]); length(are.clones)
are.traced = rownames(meta[!is.na(meta$TCR.Clone) & meta$TCR.Clone!=0, ]); length(are.traced)

# DimPlot(tils.harmony, cells.highlight=are.traced)
# DimPlot(tils.harmony, cells.highlight=are.traced, split.by='patient')

meta$patient.tcf = paste0(meta$patient, '.', meta$til.clonotype.family)
meta$patient.tcf = ifelse(meta$patient.tcf %in% c('p6.NA', 'p2.NA', 'p11.NA', 'p15.NA'), NA, meta$patient.tcf)
meta$patient.pre.tcf = paste0(meta$patient, '.', meta$til.pre.clonotype.family)
meta$patient.pre.tcf = ifelse(meta$patient.pre.tcf %in% c('p6.NA', 'p2.NA', 'p11.NA', 'p15.NA'), NA, meta$patient.pre.tcf)
meta$patient.rel.tcf = paste0(meta$patient, '.', meta$til.rel.clonotype.family)
meta$patient.rel.tcf = ifelse(meta$patient.rel.tcf %in% c('p6.NA', 'p2.NA', 'p11.NA', 'p15.NA'), NA, meta$patient.rel.tcf)

patient.tcf.freqs = meta %>% subset(!is.na(patient.tcf)) %>% group_by(patient.tcf) %>% tally(name='tcf.counts'); dim(patient.tcf.freqs)
patient.pre.tcf.freqs = meta %>% subset(!is.na(patient.pre.tcf)) %>% group_by(patient.pre.tcf) %>% tally(name='tcf.pre.counts'); dim(patient.pre.tcf.freqs)
patient.rel.tcf.freqs = meta %>% subset(!is.na(patient.rel.tcf)) %>% group_by(patient.rel.tcf) %>% tally(name='tcf.rel.counts'); dim(patient.rel.tcf.freqs)

dim(meta); head(rownames(meta))
meta$cell.barcode = rownames(meta)
# meta = m
meta = merge(meta, patient.tcf.freqs, by=c('patient.tcf'), all.x=TRUE); dim(meta)
meta = merge(meta, patient.pre.tcf.freqs, by=c('patient.pre.tcf'), all.x=TRUE); dim(meta)
meta = merge(meta, patient.rel.tcf.freqs, by=c('patient.rel.tcf'), all.x=TRUE); dim(meta)
meta[is.na(meta)]<-0


##### create a new clonotype ID so that I don't have 3 columns
meta$final.clonotype.family <- meta$patient.tcf
meta$p11.fams = paste0(meta$patient.pre.tcf, '-', meta$patient.rel.tcf)
meta$final.clonotype.family <- ifelse(meta$p11.fams!='0-0', meta$p11.fams, meta$final.clonotype.family)
final.freqs = meta %>% subset(final.clonotype.family!=0) %>% group_by(final.clonotype.family) %>% tally(name='final.family.counts'); dim(final.freqs)
meta = merge(meta, final.freqs, by='final.clonotype.family', all.x=TRUE); dim(meta)

rownames(meta)<-meta$cell.barcode
final.family = as.character(meta$final.clonotype.family); names(final.family)<-rownames(meta)
final.family.counts = as.character(meta$final.family.counts); names(final.family.counts)<-rownames(meta)
tils.harmony <- AddMetaData(tils.harmony, metadata=final.family, col.name='final.clonotype.family')
tils.harmony <- AddMetaData(tils.harmony, metadata=final.family.counts, col.name='final.family.counts')
       
```


# Find markers for each cluster
```{r top til.cd8 markers}
# find markers for every cluster compared to all remaining cells, report only the positive ones
til.markers <- FindAllMarkers(tils.harmony, min.pct = 0.25, logfc.threshold = log(2))
top_10 = til.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)

top_10 = data.frame(top_10)
DoHeatmap(tils.harmony, features = top_10$gene) + font("xy.text", size = 4)

top20 <- til.markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
DoHeatmap(tils.harmony, features = top20$gene) + font("xy.text", size = 3) 

```


#Get table clonotype distribution within clusters 
```{r echo=FALSE, eval=FALSE}
g.table = tils.harmony@meta.data[,c('orig.ident', 'sample', 'TCR.Clone', 'til.clonotype.family', 'final.clonotype.family', 'final.family.counts', 'til.pre.clonotype.family', 'til.rel.clonotype.family', 'final.clonotype.family', 'patient', 'clusters','seurat_clusters')]
g.table$cell.barcode <- rownames(tils.harmony@meta.data)
rownames(g.table) <- rownames(tils.harmony@meta.data)

```



# Just CD4 TILS
```{r subset out tils}

cells <- rownames(tils.harmony@meta.data)[(tils.harmony@meta.data$cd4.adt==1 & tils.harmony@meta.data$cd8.adt==0)]; length(cells)
tils.CD4.harmony <- subset(tils.harmony, cells = cells)
tils.CD4.harmony <- NormalizeData(tils.CD4.harmony, assay = "RNA", normalization.method = "LogNormalize", scale.factor = 1e6)

```
### Check selection
```{r plotting various things}
p1 = FeaturePlot(tils.CD4.harmony, features='cd4.adt')
p2 = FeaturePlot(tils.CD4.harmony, features='cd8.adt')
p3 = FeaturePlot(tils.CD4.harmony, features='dp.adt')
p4 = FeaturePlot(tils.CD4.harmony, features='dn.adt')
plot_grid(p1, p2, p3, p4)
DimPlot(tils.harmony, reduction='umap', label=TRUE)
DimPlot(tils.CD4.harmony, reduction='umap', label=TRUE)


```


#### Cluster again
```{r recluster}

tils.CD4.harmony <- tils.CD4.harmony %>% 
    RunUMAP(reduction = "harmony", dims=1:50) %>%
    FindNeighbors(reduction = "harmony", dims = 1:50) %>% 
    FindClusters(resolution = 0.6) %>% 
    identity()

DimPlot(tils.CD4.harmony, reduction='umap', label=TRUE)
DimPlot(tils.CD4.harmony, reduction='umap', split.by='patient')
```

#Save and reload the data
```{r save data}
tils.CD4.harmony$clusters <- tils.CD4.harmony$seurat_clusters
library(Matrix)
writeMM(obj = t(tils.CD4.harmony@assays$RNA@counts), file="cd4_counts_sparse_matrix_t.mtx")
meta = tils.CD4.harmony@meta.data[,c('seurat_clusters', 'final.family.counts', 'final.clonotype.family', 'patient', 'TCR.Clone')]
write.table(meta, 'tils_cd4_meta.txt', sep='\t', row.names=F, quote=F)
genes = data.frame('genes'=rownames(tils.CD4.harmony))
write.table(genes, 'tils_cd4_genes.txt', sep='\t', quote=F, row.names=F)

## get just cluster 4
cd4.4 = subset(tils.CD4.harmony, idents='4')
FeaturePlot(cd4.4, features=c('CD45RA.CITEseq.norm','CD62L.CITEseq.norm'), min.cutoff = 1.5)
m = cd4.4@meta.data
high = m %>% subset(CD45RA.CITEseq.norm > 1.5 & CD62L.CITEseq.norm > 1.5); dim(high)
cd4.4.high = subset(cd4.4, cells=rownames(high))
FeaturePlot(cd4.4.high, features=c('CD45RA.CITEseq.norm','CD62L.CITEseq.norm'), min.cutoff = 1.5)
rownames(high)
```



## Run UMAP on tils.CD4..harmony cells at different resolutions
```{r res til, message=FALSE, warning=FALSE}
res = c(0.30, 0.35, 0.40, 0.45, 0.50, 0.55, 0.60, 0.65, 0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1)

for (i in res) {
  tils.CD4.harmony <- FindClusters(tils.CD4.harmony, resolution=i)
  tils.CD4.harmony@meta.data[,paste0('clusters_', as.character(i))]<-tils.CD4.harmony$seurat_clusters
}

clustree(tils.CD4.harmony)

```

### Plotting different resolutions
```{r plotting tils.harmony resolutions, message=FALSE, warning=FALSE}
Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.3
d1 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.30")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.35
d2 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.35")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.4
d3 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.40")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.45
d4 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.45")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.5
d5 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.50")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.55
d6 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.55")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.6
d7 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.60")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.65
d8 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.65")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.7
d9 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.70")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.75
d10 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.75")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.8
d11 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.80")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.85
d12 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.85")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.9
d13 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.90")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_0.95
d14 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 0.95")

Idents(tils.CD4.harmony) <- tils.CD4.harmony$clusters_1
d15 = DimPlot(tils.CD4.harmony, reduction = "umap", label = TRUE) + labs(title = "Res = 1")


plot_grid(d1, d2)
plot_grid(d3, d4)
plot_grid(d5, d6)
plot_grid(d7, d8)
plot_grid(d9, d10)
plot_grid(d11,d12)
plot_grid(d15,d14, d15)

```



# Find markers for each cluster
```{r top til.cd4 markers}
# find markers for every cluster compared to all remaining cells, report only the positive ones
til.CD4.markers <- FindAllMarkers(tils.CD4.harmony, min.pct = 0.25, logfc.threshold = log(2))
top_10 = til.CD4.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)

top_10 = data.frame(top_10)
DoHeatmap(tils.CD4.harmony, features = top_10$gene) + font("xy.text", size = 4)

top20 <- til.CD4.markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
DoHeatmap(tils.CD4.harmony, features = top20$gene) + font("xy.text", size = 3) 

```



# Make Heatmaps with selected genes
```{r top til.cd4 markers}
# find markers for every cluster compared to all remaining cells, report only the positive ones
genes = c('CD4', 'CD8A','GZMK', 'GZMM', 'PRF1', 'IFNG', 'FASLG', 
             'TNF', 'CD69', 'ICOS', 'TNFRSF9', 'HLA-DRA', 'ENTPD1')

Idents(tils.CD4.harmony)<-tils.CD8.harmony$clusters
png(" ## FILE ## ", width = 10, height = 7, units='in', res=500)
DoHeatmap(tils.CD4.harmony, features = genes, slot='data', lines.width = 300) + #font("xy.text", size = 4) + 
  #theme(legend.position = "none") +
  scale_fill_gradientn(colors = c("blue", rep('chartreuse4', 20), rep('chartreuse3', 3), rep('chartreuse2', 3), rep('chartreuse1', 3), "yellow"))
dev.off()


```


#Get table clonotype distribution within clusters 
```{r echo=FALSE, eval=FALSE}
g.table = tils.CD4.harmony@meta.data[,c('orig.ident', 'sample', 'TCR.Clone', 'til.clonotype.family', 'final.clonotype.family', 'final.family.counts', 'til.pre.clonotype.family', 'til.rel.clonotype.family', 'patient', 'clusters','seurat_clusters')]
g.table$cell.barcode <- rownames(tils.CD4.harmony@meta.data)
rownames(g.table) <- rownames(tils.CD4.harmony@meta.data)


wb <- createWorkbook()
name <- "TCR clones"
addWorksheet(wb, name)
writeData(wb, sheet = name, g.table)

saveWorkbook(wb, file = 'tilsCD4.harmony.clonotype.cluster.spreadsheet.xlsx', overwrite = TRUE)

```





